---
title: "'extending ggplot2'"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scratch can be time consuming
knitr::opts_chunk$set(fig.width = 8, message = FALSE,
                      warning = FALSE, comment = "", 
                      cache = F, fig.retina = 3)
library(flipbookr)
library(tidyverse)
```

---




---

# 0. Modify existing geoms/stats

examples base on those in https://ggplot2-book.org/programming.html

---

`r chunk_reveal("existing", title = "### Modifying an existing geom")`

```{r existing, include = F}
ggplot(mpg, aes(displ, 1 / hwy)) + 
  geom_point() + 
  geom_smooth() ->
g0

geom_lm <- function(formula = y ~ x, 
                    colour = alpha("steelblue", 0.5), 
                    size = 2, ...)  {
  geom_smooth(formula = formula, 
              se = FALSE, method = "lm", 
              colour = colour,
    size = size, ...)
}

ggplot(mpg, aes(displ, 1 / hwy)) + 
  geom_point() + 
  geom_lm() ->
g1

ggplot(mpg, aes(displ, 1 / hwy)) + 
  geom_point() + 
  geom_lm(formula = y ~ poly(x, 2), 
          size = 1, colour = "red") ->
g2

library(patchwork)

g0 + g1 + g2
```


---


```{r}
tibble(num_people = c(49, 16),
       type = c("student", "professor")) %>% 
  mutate(quadrant = row_number()) %>% 
  ggplot() +
  aes(xmax = num_people^.5, ymax = num_people^.5,
      xmin = 0, ymin = 0) +
  geom_rect(color = "black") +
  coord_fixed() +
  geom_text(aes(x = num_people^.5/2,
            y = num_people^.5/2,
            label = type))

geom_sq_area <- function(area, ...)  {
  geom_rect(aes(xmax = area^.5, ymax = area^.5, 
                xmin = 0, ymin = 0),
               ...)
}

geom_sq_area_txt <- function(area, 
                    colour = alpha("steelblue", 0.5), 
                    size = 2, ...)  {
  geom_rect(aes(xmax = area^.5, ymax = area^.5, 
                xmin = 0, ymin = 0),
              colour = colour,
    size = size, ...)
}


tibble(num_people = c(49, 16),
       type = c("student", "professor")) %>% 
  mutate(quadrant = row_number()) %>% 
  ggplot() +
  geom_sq_area(area = 45,
               color = "black",
               fill = "darkolivegreen") +
  coord_equal()


tibble(num_people = c(8*8, 16),
       type = c("students", "professors")) %>% 
  ggplot() +
  aes(x = 1, y = num_people^.5, width = num_people^.5,
      fill = type) +
  geom_col(alpha = .9) +
  # coord_equal() +
  facet_wrap(~type) +
  facet_grid(~type, scales = "free_x", space = "free_x") +
  geom_text(aes(y = num_people^.5/2, 
                label = paste0(type, "\n", num_people))) +
  scale_x_continuous(expand = c(0,0)) +
  theme_void() +
  theme(legend.position = "none") +
  theme(strip.text = element_blank()) +
  theme(panel.spacing.x = unit(0, "npc"))
  

  
  

```




---

`r chunk_reveal("combining_layers", title = "### Combining existing geoms/stats into a single call")`

```{r combining_layers, include=F}
ggplot(mpg, aes(class, cty)) +
  stat_summary(fun = "mean", geom = "bar", fill = "grey70") +
  stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = 0.4) +
  geom_point(alpha = .4) ->
g0

geom_mean <- function(alpha = .4, fill = "magenta") {
  list(
    stat_summary(fun = "mean", geom = "bar", fill = fill, alpha = alpha),
    stat_summary(fun.data = "mean_cl_normal", geom = "errorbar", width = 0.4),
    geom_point(alpha = alpha)
  )
}

ggplot(mpg, aes(class, cty)) + 
  geom_mean() ->
g1

ggplot(mpg, aes(class, cty)) + 
  geom_mean(fill = "blue", alpha = .2) ->
g2

library(patchwork)
g0 + g1 + g2
```


---


# 1. Building new geoms

## a. springs

This example is https://ggplot2-book.org/spring1.html

---

`r chunk_reveal("spring", title = "### Thinking about springs")`

```{r spring, include = F}
tibble(
  x = sin(seq(0, 2*pi, length.out = 100)),
  y = cos(seq(0, 2*pi, length.out = 100)),
  index = 1:100,
  type = "circle"
) ->
  circle

spring <- circle

spring$x + seq(0, 1.5, length.out = 100) -> spring$x
spring$type <- "spring"

rbind(circle, spring) %>% 
  ggplot() + 
  facet_wrap(~ type, scales = "free_x") +
  aes(x = x, y = y, group = type, alpha = index) +
  geom_point(show.legend = FALSE) +
  geom_path(show.legend = FALSE)
```



---

`r chunk_reveal("create_spring", title = "### Writing and using a general spring function")`

```{r create_spring, include=F}
create_spring <- function(x, y, xend, yend, diameter, tension, n) {
  if (tension <= 0) {
    rlang::abort("`tension` must be larger than 0")
  }
  # Calculate direct length of segment
  length <- sqrt((x - xend)^2 + (y - yend)^2)
  
  # Figure out how many revolutions and points we need
  n_revolutions <- length / (diameter * tension)
  n_points <- n * n_revolutions
  
  # Calculate sequence of radians and x and y offset
  radians <- seq(0, n_revolutions * 2 * pi, length.out = n_points)
  x <- seq(x, xend, length.out = n_points)
  y <- seq(y, yend, length.out = n_points)
  
  # Create the new data
  data.frame(
    x = cos(radians) * diameter/2 + x,
    y = sin(radians) * diameter/2 + y
  )
}

create_spring(
  x = 4, y = 2, 
  xend = 10, yend = 6,
  diameter = 2, 
  tension = 0.75, n = 20
) %>% 
  ggplot() + 
  aes(x = x, y = y) +
  geom_point() +
  geom_path()
```

---


```{r}
StatSpring <- ggproto("StatSpring", Stat, 
  setup_params = function(data, params) {
    if (is.null(params$diameter)) {
      params$diameter <- 1
    } else if (params$diameter == 0) {
      rlang::abort("Springs cannot be defined with a diameter of 0")
    }
    if (is.null(params$tension)) {
      params$tension <- 0.75
    } else if (params$tension <= 0) {
      rlang::abort("Springs must be defined with a tension greater than 0")
    }
    if (is.null(params$n)) {
      params$n <- 50
    } else if (params$n <= 0) {
      rlang::abort("Springs must be defined with `n` greater than 0")
    }
    params
  },
  setup_data = function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  },
  compute_panel = function(data, scales, diameter = 1, tension = 0.75, 
                           n = 50) {
    cols_to_keep <- setdiff(names(data), c("x", "y", "xend", "yend"))
    springs <- lapply(seq_len(nrow(data)), function(i) {
      spring_path <- create_spring(data$x[i], data$y[i], data$xend[i], 
                                   data$yend[i], diameter, tension, n)
      cbind(spring_path, unclass(data[i, cols_to_keep]))
    })
    do.call(rbind, springs)
  },
  required_aes = c("x", "y", "xend", "yend")
)


```




---

`r chunk_reveal("unpack", title = "### Let's unpack that data manipulation that happens with a 2-row data set.")`

```{r unpack, include=F}
tibble(x = 1:2,
       y = 4:3,
       xend = 5:6,
       yend = 8:7) %>% 
  mutate(group = row_number())->
data

diameter = 1 
tension = 0.75
n = 2

setdiff(names(data), c("x", "y", "xend", "yend")) ->

  cols_to_keep

    springs <- lapply(seq_len(nrow(data)), function(i) {
      
      spring_path <- create_spring(x = data$x[i], y = data$y[i], xend = 
                                    data$xend[i], yend = 
                                   data$yend[i], diameter = diameter, 
                                    tension = tension, 
                                   n = n)
      cbind(spring_path, unclass(data[i, cols_to_keep]))
      
    })

do.call(rbind, springs) %>% 
  ggplot() +
  aes(x = x, y = y, color = group) +
  geom_point(data = data, color = "black") +
  geom_point(data = data, aes(x = xend, y = yend), color = "red") +
  geom_point()
```


---

```{r}
geom_spring <- function(mapping = NULL, data = NULL, stat = "spring", 
                        position = "identity", ..., diameter = 1, tension = 0.75, 
                        n = 50, arrow = NULL, lineend = "butt", linejoin = "round", 
                        na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomPath, 
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(
      diameter = diameter, 
      tension = tension, 
      n = n, 
      arrow = arrow, 
      lineend = lineend, 
      linejoin = linejoin, 
      na.rm = na.rm, 
      ...
    )
  )
}
```

---

`r chunk_reveal("use_spring")`

```{r use_spring, include=F}
set.seed(123)
tibble(
  x = runif(5, max = 10),
  y = runif(5, max = 10),
  xend = runif(5, max = 10),
  yend = runif(5, max = 10),
  class = 1:5
) %>% 
  ggplot() + 
  aes(x = x, y = y) +
  geom_point() +
  geom_point(aes(x = xend, y = yend), color = "red") +
  facet_wrap(~class) +
  aes(xend = xend, yend = yend) +
  geom_spring() +
  geom_spring(tension = .9, diameter = 2, 
              color = "magenta", alpha = .7)
```


---





## b. circle

This one based on the idea from T. Pedersen's talk
https://rstudio.com/resources/rstudioconf-2020/extending-your-ability-to-extend-ggplot2/

He seems to get it done in a more succinct way, so something to come back to. 


---

`r chunk_reveal("circle")`

```{r circle, include=F}
create_circle <- function(x0, y0, r = 1, n = 100){
  
  angels <- seq(
    
    from = 0,
    to = 2*pi,
    length.out = n + 1
    
  )
  
  data.frame(
    x = cos(angels) * r + x0,
    y = sin(angels) * r + y0
  )

}

create_circle(x0 = 1, y0 = 2, r = 2) %>% 
  ggplot() +
  aes(x = x, y = y) +
  geom_point() +
  geom_point(data = data_frame(x = 1, y = 2))

create_area_square <- function(area){

  data.frame(
    
    x = c(0, area^.5, area^.5, 0),
    x = c(0, 0, area^.5, area^.5)
  )
  
}

```

---




```{r}
StatCircle <- ggproto("StatCircle", Stat, 
  setup_params = function(data, params) {
    if (is.null(params$r)) {
      params$r <- 1
    } else if (params$r == 0) {
      rlang::abort("Circles cannot be defined with a radius of 0")
    }
    if (is.null(params$n)) {
      params$n <- 50
    } else if (params$n <= 0) {
      rlang::abort("Circles must be defined with `n` greater than 0")
    }
    params
  },
  setup_data = function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  },
  compute_panel = function(data, scales, r = 1, n = 50) {
    
    cols_to_keep <- setdiff(names(data), c("x0", "y0"))
    circles <- lapply(seq_len(nrow(data)), function(i) {
      circles_path <- create_circle(data$x0[i], data$y0[i], r, n)
      cbind(circles_path, unclass(data[i, cols_to_keep]))
      
    })
    do.call(rbind, circles)
  },
  required_aes = c("x0", "y0")
)
```

---



```{r}
StatCircle <- ggproto("StatCircle", Stat, 
  setup_params = function(area) {
    if (is.null(params$area)) {
      params$area <- 1
    } else if (params$r == 0) {
      rlang::abort("Square cannot be defined with an area of 0")
    }
    params
  },
  setup_data = function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  },
  compute_panel = function(data, scales, r = 1, n = 50) {
    
    cols_to_keep <- setdiff(names(data), c("x0", "y0"))
    squares <- lapply(seq_len(nrow(data)), function(i) {
      squares_path <- create_area_square(data$area[i])
      cbind(circles_path, unclass(data[i, cols_to_keep]))
      
    })
    do.call(rbind, circles)
  },
  required_aes = c("area")
)
```


---

`r chunk_reveal("circle_unpack")`

```{r circle_unpack, include=F}
tibble(x0 = 1:2,
       y0 = c(1,8),
       r = 1:2) %>% 
  mutate(group = row_number()) ->
data

r = .25
n = 10

   cols_to_keep <- setdiff(names(data), c("area"))
   
    circles <- lapply(seq_len(nrow(data)), function(i) {
      circles_path <- create_circle(data$x0[i], data$y0[i], data$r[i], n)
      cbind(circles_path, unclass(data[i, cols_to_keep]))
    })
    
do.call(rbind, circles) %>% 
  ggplot() +
  aes(x = x, y = y, col = group) +
  geom_point() +
  geom_point(data = data, aes(x = x0, y = y0))
```

```{r}

```



---


```{r}
geom_circle <- function(mapping = NULL, data = NULL, stat = "circle", 
                        position = "identity", ..., r = 1,  
                        n = 50, arrow = NULL, lineend = "butt", linejoin = "round", 
                        na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomPath, 
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(
      r = r, 
      n = n, 
      arrow = arrow, 
      lineend = lineend, 
      linejoin = linejoin, 
      na.rm = na.rm, 
      ...
    )
  )
}
```


---

`r chunk_reveal("use_circle")`

```{r use_circle, include=F}
set.seed(1244)
tibble(
  x = runif(5, max = 10),
  y = runif(5, max = 10),
  r = runif(5, max = 10),
  class = 1:5
) %>% 
  ggplot() + 
  aes(x = x, y = y) +
  geom_point() +
  aes(x0 = x, y0 = y) +
  geom_circle(n = 6) +
  geom_circle(n = 13) +
  geom_circle(r = 2) +
  geom_circle()
```

---


## c. circlemean (a collective geom)

---


### Now me... I will try to do geom_mean_circle.  This is collective geom.  Take the mean of x0 and y0 and r, then draw a circle around the mean, and r reflective of that.  

---

### collective geoms

Thus far we've looked at creating individual geoms -- a mark/geom is associated with each row in the data.  But what about a geom that summarizes.  i.e. creates a mark/geom which reflects all of the rows of data or groups within the data. This is the case of x-mean (my target) and several of the other geoms that I want to create. 

I think I'll try to get it done in the wrong way first - having a vertical line for each and every row at the mean.  And then come back to this. 

---

```{r}
StatCirclemean <- ggproto("StatCirclemean", Stat, 
  setup_params = function(data, params) {
    if (is.null(params$r)) {
      params$r <- 1
    } else if (params$r == 0) {
      rlang::abort("Circles cannot be defined with a radius of 0")
    }
    if (is.null(params$n)) {
      params$n <- 50
    } else if (params$n <= 0) {
      rlang::abort("Circles must be defined with `n` greater than 0")
    }
    params
  },
  setup_data = function(data, params) {
    if (anyDuplicated(data$group)) {
      data$group <- paste(data$group, seq_len(nrow(data)), sep = "-")
    }
    data
  },
  compute_panel = function(data, scales, r = 1, n = 50) {
   data %>% 
     summarise(x0 = mean(x0),
               y0 = mean(y0),
               r = mean(r)) %>% 
     mutate(group = 1) ->
    data
    cols_to_keep <- setdiff(names(data), c("x0", "y0"))
    circles <- lapply(seq_len(nrow(data)), function(i) {
      circles_path <- create_circle(data$x0[i], data$y0[i], r, n)
      
      cbind(circles_path, unclass(data[i, cols_to_keep]))
      
    })
    do.call(rbind, circles)
  },
  required_aes = c("x0", "y0")
)
```

---

`r chunk_reveal("circle1_unpack", title = "unpacking some data manipulation")`

```{r circle1_unpack, include=F}
tibble(x0 = 1:2,
       y0 = c(1,8),
       r = 1:2) %>%
  mutate(group = row_number()) ->
data0
data <- data0
r = .25
n = 50
   cols_to_keep <- setdiff(names(data), c("x0", "y0"))
   data %>%
     summarise(x0 = mean(x0),
                      y0 = mean(y0),
                      r = mean(r)) %>%
     mutate(group = 1) ->
    data
    circles <- lapply(seq_len(nrow(data)), function(i) {
      circles_path <- create_circle(data$x0[i], data$y0[i], data$r[i], n)
      cbind(circles_path, unclass(data[i, cols_to_keep]))
    })
do.call(rbind, circles) %>%
  ggplot() +
  aes(x = x, y = y, col = group) +
  geom_point() +
  geom_point(data = data0,
             aes(x = x0, y = y0),
             alpha = .3) +
  geom_point(data = data0,
             aes(x = x0, y = y0))
```



---


```{r}
geom_circlemean <- function(mapping = NULL, data = NULL, stat = "circlemean", 
                        position = "identity", ..., r = 1,  
                        n = 50, arrow = NULL, lineend = "butt", linejoin = "round", 
                        na.rm = FALSE, show.legend = NA, inherit.aes = TRUE) {
  layer(
    data = data, 
    mapping = mapping, 
    stat = stat, 
    geom = GeomPath, 
    position = position, 
    show.legend = show.legend, 
    inherit.aes = inherit.aes, 
    params = list(
      r = r, 
      n = n, 
      arrow = arrow, 
      lineend = lineend, 
      linejoin = linejoin, 
      na.rm = na.rm, 
      ...
    )
  )
}
```


---

`r chunk_reveal("use_circle1")`

```{r use_circle1, include=F}
set.seed(1244)
tibble(
  x = runif(5, max = 10),
  y = runif(5, max = 10),
  r = runif(5, max = 10),
  class = 1:5
) %>%
  ggplot() +
  aes(x = x, y = y) +
  geom_point() +
  geom_point(data = . %>% summarize(x = mean(x),
                                  y = mean(y)),
             color = "red") +
  aes(x0 = x, y0 = y) +
  geom_circlemean(r = 1 , n = 6) +
  geom_circlemean(r = .75 , n = 13) +
  geom_circlemean(r = .5 ) +
  geom_circle(r = .8, color = "red")
```

---



```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 65%}

@media print {
  .has-continuation {
    display: block;
  }
}
```
