---
title: "Easy Flipbook Recipes"
subtitle: "with {flipbookr} and {xaringan}"
author: "Gina Reynolds, August 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include = FALSE}
library(flipbookr)
library(tidyverse)
xaringanExtra::use_webcam()

knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F, fig.retina = 3)
```


class: inverse, center, middle

# These slides are aimed at helping you get started making flipbooks with {flipbookr} and {xaringan}.  

---
class: inverse, center, middle

# It will show you what's needed in your source .Rmd document...

---
class: inverse, center, middle

# ...and what to expect for the flipbooked result.


---
class: inverse


# In this guide:

--

##- slides with a *dark background* communicate about *'how-to'*, i.e. what will be in your .Rmd file

--

##- and slides with a *light background* demo the *flipbook experience* that you'll create upon 'knitting'. 

---
class: inverse


# I assume that you are familiar with:

##- [the R programming language](https://www.r-project.org/)
##- [knitr and Rmarkdown](https://bookdown.org/yihui/rmarkdown/) which let you comingle prose and code, and compile to documents with prose, code, output and more!

---
class: inverse

## If the {xaringan} package is new to you, things might feel a little more foreign.

--

## {xaringan} translates an Rmarkdown (.Rmd) file to an .html slide show.  

--

## For background on {xaringan} you might first check out [an introduction](https://bookdown.org/yihui/rmarkdown/xaringan.html) and later come back to a [styling guide](https://arm.rbind.io/slides/xaringan.html#1)





---
class: inverse, middle, center

# Okay. 

---
class: inverse, middle, center

# Ready to cook up some flipbooks?


---
class: inverse, middle, center

# Let's get started!

---
class: inverse

## Step 0: Installing {xaringan} and {flipbookr} 

--

- ### In your R console:

```r
install.packages("xaringan")
install.packages("remotes") # allows you to install development packages from github
remotes::install_github("EvaMaeRey/flipbookr")
```

--

- ### Note: you probably don't need to update dependency packages when installing {flipbookr}, i.e. select option #3, 'don't update' if prompted.

--

- ### Restart R.  

--

- ### Congrats! The foundation is laid for flipbooking! 



---
class: inverse, middle, center


# Step 1. Set up


---
class: inverse

## 1a. In an .Rmd file, start with a YAML like this:


````markdown
---
title: "My First Flipbook"
subtitle: "With flipbookr and xaringan"
author: "Your name here"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
````

--

### This declares your output document to be an html {xaringan} slide show.

---
class: inverse

## 1b. setup code chunk


````markdown
```{r setup, include = FALSE}`r ''`
library(flipbookr)
library(tidyverse)

knitr::opts_chunk$set(fig.width = 6, message = FALSE, 
                      warning = FALSE, comment = "", 
                      cache = F)
```
````

--

### Your setup code chunk loads {flipbookr}, and maybe other packages (like the {tidyverse} in this example) and sets some knitr defaults.



---
class: inverse

## 1c. css code chunk

````markdown
```{css, eval = TRUE, echo = FALSE}`r ''`
.remark-code{line-height: 1.5; font-size: 80%}

@media print {
  .has-continuation {
    display: block;
  }
}
```
````

--

### css stands for cascading style sheets (you'll see that in the YAML too), and this code chunk helps define the look of the html document you'll ultimately produce. 

--

### In this chunk, you can quickly adjust the font size for the code that you'll be revealing.  Now it is set to 80%.

---
class: inverse, middle, center

# Step 2: building a code pipeline in a 'source' code chunk

---
class: inverse

### A *source* code chunk, named 'my_cars' will serve as the material to be flipbooked. 

````markdown
```{r task1, include = FALSE}`r ''`
gapminder %>% 
  filter(year == 2007) %>% 
  select(country, lifeExp, year) %>% 
  filter(lifeExp >= 80) %>% 
  arrange(-lifeExp)
```
````

--

#### We'll refer to this code chunk later.  We don't want to see this chunk itself, so you should set `include = FALSE`


---
class: inverse, middle, center

# Step 3: Ask for flipbooked version of source code

---
class: inverse


### In your .Rmd use the {flipbookr} function `chunk_reveal()` *inline* as follows.  Refer to the source code chunk that you have prepared by name. 

```markdown
---

``r 'r chunk_reveal(chunk_name = "my_cars", title = "### First flipbook!")'``
``` 


--

### Also, indicate you want a slide break before the inline code  as shown above. The slide break is indicated with three dashes, '---' at the beginning of a line with no trailing spaces.

--

### Let's see the result... (you might 'knit' your .Rmd now)


---

```{r my_cars, include=FALSE}
cars %>% 
  ggplot() +
  aes(x = speed) +
  aes(y = dist) + #BREAK
  geom_point(
             shape = 21,
             size = 7, #BREAK2
             color = "magenta", #BREAK3
             alpha = .8 #BREAK4
             ) + #BREAK
  aes(fill = speed) + 
  scale_fill_viridis_c() + 
  theme_minimal() #BREAK
```

`r chunk_reveal("my_cars", title = "### First flipbook!")`




---
class: inverse, center, middle

# Step 4: Modify the recipe

---
class: inverse

## We've looked at the *classic* flipbook recipe -- what `chunk_reveal()` gives us when basically use *defaults*.

--

## The classic flipbook is a side-by-side *code-output* reveal, where break points are determined *automatically*.  `chunk_reveal()` pauses where parentheses are balanced at ends of lines.  


---
class: inverse

## Let's change the recipe. 

--

### We get different build behavior by setting the `break_type` argument to `"user"`.  We can refering to the same chunk, 'my_cars'.  

```markdown
---

``r 'r chunk_reveal("my_cars", break_type = "user", title = "### Another code movie!")'``
``` 

--

### When break_type = "user", flipbookr's `chunk_reveal()` looks for the special comment *#BREAK* in the source code chunk, and displays code sequentially at the *user defined* breaks.

--

### Let's see the result... (you might 'knit' your .Rmd now)

---

`r chunk_reveal("my_cars", break_type = "user", title = "### Another code movie!")`


---
class: inverse

### Let's change the recipe again.  Setting the `break_type` argument to `"non_seq"` allows you to define a non sequential order of display in the flipbook. 


```markdown
---

``r 'r chunk_reveal("my_cars", break_type = "non_seq")'``
``` 

--

### Here the comments #BREAK2, #BREAK3 etc. in the 'my_cars' code chunk inform flipbookr of which lines of code should be displayed in which frames of the flipbook. Lines without such a comment are always displayed.


--

### Let's see the result ...

---

`r chunk_reveal("my_cars", break_type = "non_seq")`

---
class: inverse

## The final recipe we'll look at is changing the `display_type` from its default, which is `c("code", "output")`.

---
class: inverse


### Let's build a *new* source code chunk, "my_mtcars".

````markdown
```{r my_mtcars, include = F}`r ''`
mtcars %>% 
  ggplot() +
  aes(x = factor(cyl)) +
  aes(y = mpg) +
  geom_boxplot() +
  geom_jitter(alpha = .5,
              width = .25)
```
````


```{r my_mtcars, include = F}
mtcars %>% 
  ggplot() +
  aes(x = factor(cyl)) +
  aes(y = mpg) +
  theme_gray(base_size = 20) +
  geom_boxplot() +
  geom_jitter(alpha = .5,
              width = .25)
```

--

#### Then we use `chunk_reveal()` on this the "my_mtcars" code chunk and set display_type = "output" to give us *only the output panel*. 

```markdown
---

``r 'r chunk_reveal("my_mtcars", display_type = "output")'``
```


--

### Let's look at the flipbook ...

---

`r chunk_reveal("my_mtcars", display_type = "output")`




---
class: inverse, 

# Congratulations! 

--

### Now your are cooking with {flipbookr} and {xaringan}. We went through these steps:

###- Step 0: Install {flipbookr} and {xaringan}
###- Step 1: Set up .Rmd (a. YAML, b. setup code chunk, c. css code chunk)
###- Step 2: Write a source code to be flipbooked
###- Step 3: Use {flipbookr}'s chunk_reveal to build flipbook (upon knit)


---
class: inverse, 

## *And* you have learned *four flipbook recipes*!

--

###1. *The classic* (defaults code-output reveal at auto break points)
--

###2. *The picky eater* **break_type = "user", with #BREAK comments in source code chunk** user defines break points

--

###3. *Tossed salad* **break_type = "non_seq", with #BREAK2, #BREAK3 comments in source chunk** user sets order of break points, which may be non sequential

--

###2. *Hold the code* **display_type = "output"** display only an output panel





---
class: inverse

# Afterward: Source for a tiny flipbook and output

## You might find it useful to at an .Rmd for a full (but tiny) slide show with flipbooks and the resultant html.

##- [.Rmd](https://raw.githubusercontent.com/EvaMaeRey/flipbooks/master/tiny_flipbook.Rmd)

##- [.html](https://evamaerey.github.io/flipbooks/tiny_flipbook.html)



---
class: inverse

## Epilogue: Share your work and learn more recipes


### Once you master flipbooking you'll probably want to share your work

--

### But the .html slides that {xaringan} produces aren't contained in a single file, so doing so can be a little tricky! 

--

### Instead a folder of associated files will be created, and this needs to travel with your work.  

---
class: inverse

## Epilogue: Share your work and learn more recipes

### For more about ways to share your slide shows and for more flipbook recipes consult the template that travels with the {flipbookr} package.  

--

### In RStudio access this .Rmd template via: File -> New File -> RMarkdown -> From Template -> A Minimal Flipbook

--

## You may check out the template rendered to html [here](https://evamaerey.github.io/flipbooks/flipbookr/skeleton.html).


---
class: inverse

# Acknowledgments

### Emi Tanaka and Garrick Aden-Buie were critical in getting this project off the ground and have provided much help, support and feedback in the development.

### Tidyverse tools, and especially ggplot2, have been an inspiration for flipbooks. 

### The knitr/Rmarkdown ecosystem are {xaringan} are of course a key infrastructures this project relies on.

---
class: inverse

# Spread the word and giving feedback

### Please help us spread the word about flipbooks.  Consider a quick acknowledgment like: *The 'flipbooked' portion of this presentation was created with the new {flipbookr} package.  Get it at remotes::install_github("EvaMaeRey/flipbookr")*


### Feedback? Contributions? Leave an issue at:  https://github.com/EvaMaeRey/flipbookr 


