---
title: "flipbookr building blocks"
subtitle: "what lies beneath"
author: "Gina Reynolds, December 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["xaringan-themer.css", ninjutsu]
    nature:
      ratio: 30:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup}
library(flipbookr)
library(tidyverse)
knitr::opts_chunk$set(fig.width = 6, message = F, warning = F, comment = "", cache = F)
```

```{r xaringan_look}
xaringanthemer::mono_accent(
  code_highlight_color = "rgba(255, 211, 0, .35)",
  base_color = "#03418A")
```

---


We'll use the tidyverse packages to create a pipeline, and flipbookr to walk through the steps in this pipeline.

```{r}
library(tidyverse)
library(flipbookr)
```

The input code to be "flipbookified" is below:

```{r cars_plot, echo = T, eval = F, fig.height=5}
cars %>%  #BREAK          
  ggplot() +  
  aes(x = speed) + 
  aes(color = speed) + #BREAK2
  geom_rug() #BREAK
```


---

Let's first look at the result of using flipbookr to bring this code.

---

`r chunk_reveal("cars_plot")`

---


# Move to automate


## 1. *Automatically* parsing a block of input code 

--

## 2. *Automatically* reconstructing code into partial builds

--

## 3. *Automatically* delivered on a presentation platform (here Xaringan)



---

Now, lets look at some of the functions working in the background.


---
class: middle center inverse

# Step 1
## Parse block of input code


---




---
class: middle center inverse

# Step 2
## Reconstruct code into partial builds

---

## First calculate which lines need to be shown and highlighted in each frame of the mini flipbook.

--

A list is returned where each frame corresponds to the element in the list. 

--

Within the list numbers of the lines of code for each frame is returned.

--

This is dependent on what the user has set for the break_type. 

--

Appropriate highlighting is calculated based on what appears in previous frame and what is new in current frame. 


---


`r chunk_reveal("the_code_parsing", width_left = "32%", width_middle = "32%", width_right = "32%",  display_type = c("code", "func", "output"))`

```{r the_code_parsing, include=FALSE}
library(flipbookr)

create_code() %>%
  code_as_table() %>%
  code_as_table_process_break_messages() ->
hide_away

create_code() %>%
  r_code_base_parse() %>% 
  r_base_parsed_count_parentheses() ->
hide_again


create_code() %>% 
  r_code_full_parse() ->
hide_more

create_code() %>% 
  code_parse() %>% 
  parsed_calc_show() %>% 
  shown_lines_calc_highlight() ->
hiding
```

---

## Based on the calculation, return partial code builds for each frame of the flipbook.

--

We create a list of strings of the partial builds, which can be saved as an object `code_seq`.

---


`r chunk_reveal("return_partial_and_sequence", width_left = "32%", width_middle = "32%", width_right = "32%",  display_type = c("code", "func", "output"), break_type = "rotate")`

```{r return_partial_and_sequence, eval = F, echo = F} 
create_code() %>% 
  code_parse() %>% 
  parsed_return_partial_code() #ROTATE
  parsed_left_assign_return_partial_code() #ROTATE
  parsed_return_partial_code_sequence() #ROTATE
  parsed_return_recent_function() #ROTATE
  parsed_return_recent_function_sequence() #ROTATE
```




---
class: middle center inverse

# Step 3 
## Deliver to a presentation platform (here Xaringan)

---

## We have generic code chunk templates ...

---

`r chunk_reveal("returns", width_left = "32%", width_middle = "32%", width_right = "32%",  display_type = c("code", "func", "output"), break_type = "rotate")`

```{r returns, eval = F, echo = F} 
return_partial_chunks_template_code() #ROTATE
return_partial_chunks_template_output() #ROTATE
return_partial_chunks_template_function() #ROTATE
return_markdown(text = "my. short. text.", sep = "\\.") #ROTATE

```



---

## ... which are used to create chunks of code and output displayed side-by-side

```{r}
chunk_expand(chunk_name = "cars_plot")
```


---

## ... or a series of slides that *just* display output or *just* display code 


```{r}
chunk_expand(num_breaks = 2, display_type = "code")
chunk_expand(num_breaks = 2, display_type = "output")
```




---

# The reveal asks for the chunks we've created to be evaluated, by using the knitr::knit() function. 

We've applied it above.  Let's just look at the function itself.  



---

`r chunk_reveal("last", width_left = "32%", width_middle = "32%", width_right = "32%",  display_type = c("code", "func", "output"), break_type = "rotate")`

```{r last, eval = F, echo = F} 
define_css() #ROTATE
chunk_expand() #ROTATE
chunk_reveal #ROTATE
```



---


# Go to code [**here**](https://github.com/EvaMaeRey/flipbookr/blob/master/R/base_parse_reveal_xaringan.R)

https://github.com/EvaMaeRey/flipbookr/blob/master/R/base_parse_reveal_xaringan.R





```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 60%}
```