---
title: "A getting started with {flipbookr}"
author: "Gina Reynolds, August 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---


```{r setup, include = FALSE}
library(flipbookr)

knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = F, fig.retina = 3)
```


class: inverse, center, middle



# This set of slides is aimed at helping you get started with {flipbookr}.  It will show you the text that you'll use in your source .Rmd document and will show you the flipbook result.


---
class: inverse


# In this guide:

--

##- slides with a *dark background* communicate about *'how-to'*, i.e. what will be in your .Rmd file

--

##- and slides with a *light background* demo the *flipbook experience* that you'll create.

---
class: inverse, middle, center


# Set up

--

#### (The template will give you this too once you have installed flipbookr.  File -> New File -> RMarkdown -> From Template -> A Minimal Flipbook  But let's have a look.)

---
class: inverse

### YAML metadata will look like this 



````markdown
---
title: "How to Flipbook"
subtitle: "With flipbookr and xaringan"
author: "Gina Reynolds, December 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
````


---
class: inverse

## setup

#### Your set up code chunk loads {flipbookr} and sets some knitr defaults

````markdown
```{r setup, include = FALSE}`r ''`
library(flipbookr)

knitr::opts_chunk$set(fig.width = 6, message = FALSE, 
                      warning = FALSE, comment = "", 
                      cache = F, dev = "svg")
```
````

---
class: inverse

## A source code chunk

#### We'll refer to this code chunk later.  We don't want to see this chunk itself, so, you can set `include = FALSE`

````markdown
```{r cars, include = FALSE}`r ''`
library(tidyverse)

cars %>% 
  ggplot() +
  aes(x = speed) +
  aes(y = dist) +
  geom_point(shape = 21,
             size = 7, #BREAK2
             color = "white", #BREAK3
             alpha = .8 #BREAK4
             ) +
  aes(fill = speed) +
  scale_fill_viridis_c() +
  theme_minimal()
```
````

---
class: inverse




### In your .Rmd:

```markdown
``r 'r chunk_reveal(chunk_name = "cars", title = "## Our first flipbook!")'``
``` 





---


`r chunk_reveal("cars", title = "## Our first flipbook!")`

```{r cars, include=FALSE}
library(tidyverse)

cars %>% 
  ggplot() +
  aes(x = speed) +
  aes(y = dist) + #BREAK
  geom_point(shape = 21,
             size = 7, #BREAK2
             color = "white", #BREAK3
             alpha = .8 #BREAK4
             ) + #BREAK
  aes(fill = speed) + #BREAK
  scale_fill_viridis_c() + #BREAK
  theme_minimal() #BREAK
```



---
class: inverse

## Refer to same chunk, 'cars', get different break behavior

### In your .Rmd:

```markdown
``r 'r chunk_reveal("cars", break_type = "non_seq", title = "## user defined reveal (before we saw auto)")'``
``` 

---

`r chunk_reveal("cars", break_type = "non_seq", title = "## user defined reveal (before we saw auto)")`


