---
title: "Racing Barchart with gganimate"
subtitle: "a quick flipbook | made with Xaringan"
author: "<br><br>Gina Reynolds, April 2019"
date: "<br>"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.height = 6, out.width = "100%", comment = " ", cache = T, message = F, warning = F, dpi = 300)

source("https://raw.githubusercontent.com/EvaMaeRey/little_flipbooks_library/master/xaringan_reveal_parentheses_balanced.R")
```


# Where we are going: 

We'll create this checkerboard (something Andrew Heiss was recently working on).  I have **echo** set to FALSE in the code chunk options here so that you don't see the code, and **eval** to TRUE so that the plot output is produced.  The code chunk is given the name "checkerboard", and this is used in the in apply_reveal function, which will break down the code by wrangling and plot statements. 

```{r checkerboard, echo = F, eval = T, fig.height=5}
crossing(x = 1:8, y = 1:8) %>%
  mutate(value = (x + y) %% 2) ->
my_data
```


```{r my_plot}
  ggplot(data = my_data) + 
  aes(x = x, 
      y = y, 
      fill = factor(value)) +
  geom_tile() +
  theme_void() +
  scale_fill_manual(
    values = c("thistle", 
               "darkgrey"), 
    guide = F) + 
  coord_equal()
```
 

 
---

# How do we get there?

In the next slide, we'll walk through the code that produces this plot, and the output along the way. We use the code **apply_reveal("checkerboard")** in-line to access the code from the code chunk called *checkerboard*.
 

```{r}
reveal(chunk_name = "my_plot", 5, 5)
```




```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 60%}
```










