---
title: "tuning ggplot themes"
subtitle: "a flipbook | made with Xaringan"
author: "<br><br>Gina Reynolds, May 2019"
date: "<br>"
output:
  xaringan::moon_reader:
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css: ["kunoichi", "ninjutsu"]
    nature:
      ratio: "16:9"
      highlightLines: true
---


```{r}
library(tidyverse)
source("../xaringan_reveal_parentheses_balanced.R")
```



```{r example}
cars %>%            
  mutate(speed_14_plus = speed >= 14) %>%  
  ggplot() +              
  aes(x = speed) +
  aes(y = dist) +  
  # Describing what follows
  geom_point(
    size = 8, #REVEAL5
    aes(color = speed_14_plus, #REVEAL6
        shape = speed_14_plus), #REVEAL6
    ) +
  geom_point(
    data = cars %>% filter(speed > 14), #REVEAL2
    color = "red", #REVEAL2
    size = 5,   #REVEAL3
    alpha = .8, #REVEAL4
    )
```

```{r}
"example" %>% 
  parse_chunk() %>% 
  knitr::kable()
```



```{r}
local_code <- "cars %>%            
  mutate(speed_14_plus = speed >= 14) %>%  
  ggplot() +              
    aes(x = speed) +
    aes(y = dist) +  
    # Describing what follows
    geom_point(
      size = 2,   #REVEAL2
      alpha = .3, #REVEAL3
      color = \"blue\", #REVEAL4
      ) + 
    aes(color = speed_14_plus)"
```


---

```{r}
parse_code(local_code)
parsed <- parse_code(local_code)

reveal_parsed_nonsequential <- function(parsed, which_supress = 2:5, which_highlight = 6){
  
  parsed %>% 
    mutate(reveal = ifelse(row_number() %in% which_supress, "", raw_code)) %>% 
    mutate(highlight = ifelse(row_number() %in% which_highlight, "#<<", "")) %>% 
    mutate(out = paste0(reveal, "  ", comment, highlight)) %>% 
    select(out) ->
    up_to_result
  up_to_result$out  
  
}


reveal_code_nonsequential <- function(code, which_supress = 2:5, which_highlight = 6){
  
  parsed <- parse_code(code = code) 

  parsed %>% 
    mutate(reveal = ifelse(row_number() %in% which_supress, "", raw_code)) %>% 
    mutate(highlight = ifelse(row_number() %in% which_highlight, "#<<", "")) %>% 
    mutate(out = paste0(reveal, "  ", comment, highlight)) %>% 
    select(out) ->
    up_to_result
  up_to_result$out  
  
}


reveal_chunk_nonsequential <- function(chunk_name, which_supress = 2:5, which_highlight = 6){
  
  content <- chunk_as_text(chunk_name)
  parsed <- parse_code(code = content) 

  parsed %>% 
    mutate(reveal = ifelse(row_number() %in% which_supress, "", raw_code)) %>% 
    mutate(highlight = ifelse(row_number() %in% which_highlight, "#<<", "")) %>% 
    mutate(out = paste0(reveal, "  ", comment, highlight)) %>% 
    select(out) ->
    up_to_result
  up_to_result$out  
  
}


local_code %>% 
  parse_code() %>% 
  reveal_parsed_nonsequential() 
```


```{r}
reveal_chunk_nonsequential('example', 1, NULL)
reveal_chunk_nonsequential("example")
```


```{r}
calc_supress <- function(parsed){
  
  parsed %>% 
    pull(user_non_seq) ->
  code_ordering  
    
  code_ordering %>% 
    unique() %>% 
    sort() ->
  steps
  
  which_supress <- list()
  
  for (i in steps) {
    
    which_supress[[i]] <- which(code_ordering > i)
      
  }

  
  return(which_supress)
  
}


local_code %>% 
  parse_code() %>% 
  calc_supress()

calc_highlight_non_sequential <- function(parsed) {
  
  parsed %>% 
    pull(user_non_seq) ->
  code_ordering  
    
  code_ordering %>% 
    unique() %>% 
    sort() ->
  steps
  
  highlighting <- list()
  
  for (i in steps)
  
  if (i == 1) {
    highlighting[[i]] <- NULL
  } else {
    highlighting[[i]] <- which(code_ordering == i)
  }
  
  return(highlighting)

}

"example" %>% 
  parse_chunk() %>% 
  calc_highlight_non_sequential()

```


```{r}
# partial knit chunks

non_sequential_partially_knit_chunks <- function(chunk_name, show_code = T, title = "", reg_assignment = F) {
  # Create slide for lines 1:N for each line N in the given chunk
  
  parsed <- parse_chunk(chunk_name)
  
  parsed %>% 
    pull(user_non_seq) %>% 
    unique() %>% 
    sort() ->
  steps
    
  steps_supress <- calc_supress(parsed)

  highlighting <- calc_highlight_non_sequential(parsed)
  
  if (show_code == T) {
    partial_knit_steps <- glue::glue(
      "class: split-40",
      "count: false",
      ".column[.content[",
      "```{r plot_{{chunk_name}}_{{steps}}, eval=FALSE, code=reveal_chunk_nonsequential('{{chunk_name}}', {{steps_supress}}, {{highlighting}})}",
      "```",
      "]]",
      ".column[.content[",
      "```{r output_{{chunk_name}}_{{steps}}, echo=FALSE, code=reveal_chunk_nonsequential('{{chunk_name}}', {{steps_supress}}, {{highlighting}})}",
      "```",
      "]]",
      .open = "{{", .close = "}}", .sep = "\n"
    )
  } else {
    
    partial_knit_steps <- glue::glue(title,"```{r output_{{chunk_name}}_{{steps}}, echo=FALSE, code=reveal_chunk_nonsequential('{{chunk_name}}', {{steps_supress}}, {{highlighting}})}",
                                     "```",
                                     .open = "{{", .close = "}}", .sep = "\n"
    )
    
  }
  
  glue::glue_collapse(x = partial_knit_steps, sep = "\n---\n")
}


```

---

```{r}
non_sequential_partially_knit_chunks(chunk_name = "example")
```

---

```{r}
apply_reveal_nonsequential <- function(...){
  paste(knitr::knit(text = non_sequential_partially_knit_chunks(...)), collapse = "\n")
}
```

---

`r apply_reveal_nonsequential("example")`



---



```{r caution_group, eval = F, echo = F}
library(gganimate)
library(gapminder)
gapminder %>% 
  sample_frac(size = 1, replace = F) %>% #REVEAL2
  ggplot() +
  aes(x = gdpPercap) + 
  aes(y = lifeExp) +
  aes(size = pop) + 
  aes(colour = continent) +
  aes(group = country) + #REVEAL3
  geom_point(alpha = 0.7, 
             show.legend = FALSE) +
  scale_size(range = c(2, 12)) +
  scale_x_log10() +
  facet_wrap(~ continent) +
  # Here comes the gganimate specific bits
  labs(title = 'Year: {frame_time}', 
       x = 'GDP per capita', 
       y = 'life expectancy') +
  gganimate::transition_time(year) +
  gganimate::ease_aes('linear')
```

`r apply_reveal_nonsequential("caution_group")`


```{css, eval = TRUE, echo = F}
.remark-code{line-height: 1.5; font-size: 80%}
```

