---
title: "Measles Data"
subtitle: "With flipbookr and xaringan"
author: "Gina Reynolds, December 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = T, fig.retina = 3)
library(flipbookr)
library(tidyverse)
```

---

This book uses measles data from the Wall Street Journal.  I don't totally trust the variable that I'm mapping with is mmr -- it is supposed to be Measles, Mumps and Rubella vaccinations.  There are a lot of schools where the % vaccinated is zero.  This seems implausible, but I haven't hunted down the reason or what zero might mean. So the graph just take the data at face value, which is probably wrong. 


---

# tidytuesday's cleaning

The cleaning steps are from TidyTuesday.  


---

`r chunk_reveal("clean")`

```{r clean, include = F}
library(tidyverse)
library(rvest)

url_wsj <- "https://raw.githubusercontent.com/WSJ/measles-data/master/all-measles-rates.csv"

read_csv(url_wsj) ->
wsj_raw

list_of_urls <- "https://github.com/WSJ/measles-data/tree/master/individual-states"

list_of_urls %>% 
  read_html() %>% 
  html_table() %>% 
  .[[1]] %>% 
  select(Name) %>% 
  mutate(Name = str_remove(Name, 
                           "\\.csv")) %>% 
  filter(str_length(Name) > 3, 
         str_length(Name) < 20) %>% 
  pull(Name) ->
raw_states

glue::glue("https://raw.githubusercontent.com/WSJ/measles-data/master/individual-states/{raw_states}.csv") %>% 
  map(read_csv) ->
all_states

all_states %>% 
  purrr::map(~select(., state, name, lat, lng)) %>% 
  purrr::map(~mutate_at(., vars(lat, lng), as.numeric)) %>% 
  bind_rows() %>% 
  filter(!is.na(lat)) ->
clean_states

wsj_raw %>% 
  left_join(clean_states, by = c("name", "state")) ->
measles
```

---


```{r}
measles %>% 
 ggplot() +
  labs(title = "School's Measles, Mumps, and Rubella (MMR) vaccination rate, 2018-19") +
  aes(x = mmr) +
  geom_histogram()
```


---

`r chunk_reveal("plot")`

```{r plot, include = F}
measles %>% 
  filter(year == "2018-19") %>%
  filter(lng < 0) %>% 
  mutate(mmr_range = case_when(mmr < 15 ~ "low",
                               mmr >= 15 & mmr < 85 ~ " medium",
                               mmr >= 85 ~ "  high")) %>% 
  ggplot() +
  labs(title = "School's Measles, Mumps, and Rubella (MMR) vaccination rate, 2018-19") +
  aes(x = lng) +
  aes(y = lat) +
  geom_point(alpha = .02) +
  aes(color = mmr) +
  theme_minimal() +
  coord_map(projection = "mollweide") +
  scale_colour_viridis_c(direction = -1, end = .7) +
  facet_wrap(~mmr_range, nrow = 2) +
  geom_point(data = . %>% 
               select(-mmr_range) ,
             color = "grey") +
  geom_point(alpha = .02, 
             color = "magenta") +
  theme(legend.position = "none") +
  theme(axis.text.x = element_blank()) +
  theme(axis.text.y = element_blank()) +
  labs(y = NULL) +
  labs(x = NULL) 
```




---

```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 60%}
```
