---
title: "Cheers for Flipbooks"
output:
  html_document:
    # df_print: paged
  html_notebook:
    nature:
      beforeInit: https://platform.twitter.com/widgets.js
---




```{r, include=F}
knitr::opts_chunk$set(cache = F, echo = F, warning = F, message = F, error = F)
```



```{r, results='asis'}
library(tidyverse)
load("data_products/tweets.Rdata")
for (i in 1:nrow(df)) {
  cat(paste(df$lines[i], "\n"))
  cat(df$id[i])
}
```




Additionally, it has been linked to from other resources including blogs:

https://huygens.science.uva.nl/ggPlotteR/
https://ggplot2tutor.com/howitworks
https://bookdown.org/martin_monkman/DataScienceResources_book/datavisualization.html
https://medium.com/open-data-delaware/moving-outside-the-box-for-data-visualization-efcda44ce7ec
https://datascience.blog.wzb.eu/2019/03/01/linkdump-105/
https://bookdown.org/martin_monkman/DataScienceResources_book/datavisualization.html
https://community.plantae.org/files/posts/5255016927754977602/9be5f23ce0e1a9d173c263e6465b9ef9_DataVisMagda.pdf
http://bryklab.net/compbioftw/
http://stefanorodighiero.net/links.html
https://www.facebook.com/1673798376000921/posts/2035217649858990/
https://tidytuesday.fireside.fm/2
https://bookdown.org/martin_monkman/DataScienceResources_book/datavisualization.html
https://rworkshop.uni.lu/lectures/lecture06_plotting.html#46
http://mribeirodantas.xyz/blog/index.php/tag/biology/
https://www.williamrchase.com/slides/ggplot_intro.html#1
https://github.com/zhiiiyang/XaringanTutorial
https://libguides.kaust.edu.sa/ReproducibilityWS/materials
https://threadreaderapp.com/thread/1100663931192963072.html
https://bookdown.org/oscar_baruffa/Book-Twitter-for-R/follow-some-more-folks.html
https://stackoverflow.com/questions/55912039/building-up-a-ggplot-one-line-at-a-time-for-a-presentation
http://www.rpubs.com/dgolicher/ggplots
https://www.quora.com/What-data-visualization-software-can-handle-large-amounts-of-data-in-the-order-of-hundreds-of-millions-I-have-had-trouble-with-Tableau


{flipbookr}, the package that I've built for easily creating flipbooks is being used/explored by:

- [The Hirschey Lab](https://github.com/matthewhirschey/tidybiology-meetup/blob/7d0a51fb65e25615ad65dae49a92980a5b382096/2020_02/code/prism_plot_flipbook.Rmd) - The Hirschey Lab studies different aspects of metabolic control, mitochondrial signaling, and cellular processes regulating human health and disease
- [OCTRI Biostatistics, Epidemiology, Research & Design - Jessica Minnier](https://jminnier-berd-r-courses.netlify.com/04-ggplot/04_ggplot_slides.html#1) - Assistant Professor of Biostatistics, OHSU
- [Data Wrangling and Visualisation in R Oxford IT Services - Justin Miller](https://justinmillar.github.io/ox-stats-r-workshop/#1) - Post-doc researcher with the Malaria Atlas Project
- [R Course for INDEC (Instituto Nacional de Estadistica y Census Republica Argentina)](https://github.com/pablotis/Curso_R_INDEC/blob/a6dd8eadff200ceaac13e8f8d548aed8c257f243/Clase%204/Clase_4_Presentacion.Rmd)
- [Derek Ogle, Author of fishR: R for fisheries analysis](https://github.com/EvaMaeRey/flipbookr/issues/17)

<!-- "First, thank you very much for flipbookr. It looks to be a wonderful tool!!" - Professor at Northland College, Author of Introductory Fisheries Analyses with R, Maintainer of fishR -->
