---
title: "`aes` is pronounced 'ask for representation'"
subtitle: "ggplot2::aes()"
author: "Gina Reynolds, March 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = T, fig.retina = 3)
library(flipbookr)
library(tidyverse)
```



---
class: center, middle


Let's see how we get this representation using ggplot2.  First we'll do a walk through in a quirky way.  Then look at the same plot build with the normal syntax.  



---

`r chunk_reveal("short_for")`

```{r, short_for, include = F}
ASK_FOR_REPRESENTATION <- aes
CHANGE_FROM_DEFAULT_LABEL <- labs
gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  ggplot() +
  ASK_FOR_REPRESENTATION(x = gdpPercap) +
  ASK_FOR_REPRESENTATION(y = lifeExp) +
  geom_point() +
  ASK_FOR_REPRESENTATION(color = continent) +
  ASK_FOR_REPRESENTATION(shape = continent) +
  ASK_FOR_REPRESENTATION(size = pop) +
  CHANGE_FROM_DEFAULT_LABEL(x = "GDP Per cap") +
  CHANGE_FROM_DEFAULT_LABEL(y = "Life Expectancy") +
  CHANGE_FROM_DEFAULT_LABEL(size = "Population") +
  CHANGE_FROM_DEFAULT_LABEL(color = "") +
  CHANGE_FROM_DEFAULT_LABEL(shape = "") +
  CHANGE_FROM_DEFAULT_LABEL(title = "Life Expectancy v. GDP per Cap in 2002") +
  CHANGE_FROM_DEFAULT_LABEL(subtitle = "Data source: gapminder package in R") +
  CHANGE_FROM_DEFAULT_LABEL(caption = "Viz: Gina Reynolds @EvaMaeRey")
```

---
class: middle, center

*Now the normal ggplot2 syntax*

---

`r chunk_reveal("normal")`

```{r, normal, include = F}
# the normal way ...
gapminder::gapminder %>% 
  filter(year == 2002) %>% 
  ggplot() +
  aes(x = gdpPercap) +
  aes(y = lifeExp) +
  geom_point() +
  aes(color = continent) +
  aes(size = pop) +
  labs(x = "GDP Per cap") +
  labs(y = "Life Expectancy") +
  labs(size = "Population") +
  labs(color = "") +
  labs(title = "Life Expectancy v. GDP per Cap in 2002") +
  labs(subtitle = "Data source: gapminder package in R") +
  labs(caption = "Viz: Gina Reynolds @EvaMaeRey")
```

---
class: centered, middle

*The flipbooked portion of this presentation was created with the new {flipbookr} package.  Info: https://github.com/EvaMaeRey/flipbookr* 




---

r chunk_reveal("what_representation", break_type = "user", display_type = "output")`

```{r what_representation, include = F}
tibble(num_id = 1:10, id = c("Alicia", "Brian", "Candice", "Doug", "Emily", 
              "Frank", "Gary", "Harry", "Ingrid", "Jasmine"), height = round(rnorm(10), 1) + 10) %>% 
  ggplot() +
  labs(title = "What aesthetics can do \nrepresentative work for us?") +
  theme_void(base_size = 15) +
  aes(x = height) +
  aes(y = 0) +
  labs(subtitle = "x position") + 
  aes(label = paste(id, height, sep = ":")) +
  ggrepel::geom_text_repel(force = 4, 
                           color = "black",
                           size = 3,
                           alpha = 1) ->
base
base +
  geom_point(shape = "|",
             size = 2) + #BREAK
  labs(title = NULL) -> 
x_position_aes
base +   
  aes(y = height, x = 0) +
  labs(subtitle = "y position") +
  geom_point(shape = "_",
             size = 2) + #BREAK
  labs(title = NULL) -> 
y_position_aes
base +
  aes(x = 0, y = num_id) +
  scale_y_reverse() +
  aes(fill = height) +
  labs(subtitle = "color") +
  scale_fill_viridis_c() + 
  scale_x_continuous(expand = c(1,1)) +
  geom_tile(width = .2, height = .8,
            alpha = .7) + #BREAK
  labs(title = NULL) -> 
color_aes
base +
  aes(x = 0, y = num_id) +
  scale_y_reverse() +
  geom_point(shape = "diamond") +
  aes(size = height) + #BREAK
  labs(title = NULL) -> 
size_aes
base +
  aes(x = 0, y = id) +
  scale_y_reverse() +
  labs(subtitle = "size") + 
  aes(size = NULL) +
  aes(alpha = height) +
  geom_tile(width = .2, height = .8,
            alpha = .7) +
  labs(subtitle = "transparency") + #BREAK
  labs(title = NULL) -> 
transparency_aes 
```

---






```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 80%}
```
