---
title: "Mapping Data in R with geom_sf()"
subtitle: "A 'flipbookr' Tutorial"
author: "Conner J. Surrency, January 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 10:5
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, echo=F}
knitr::opts_chunk$set(fig.width = 12, message = FALSE, warning = FALSE, comment = "", cache = F, fig.retina = 3)
library(flipbookr)
```

# What Are We Doing?
--

- Here we will walk through the process of creating maps in RStudio.
--

- This will include the packages required and how to code a map.
--

- We will be looking at data from two cases: Wetherspoon restaurants in UK, and defense expenditure of NATO member states. 


---

## Set-Up
--

We will need the following packages (and data) installed:  
--

- tidyverse
--

- rnaturalearth


---
#Example #1
--

We know the location of restaurants in the UK and the price of the empire burgur. 


The map we make will be a simple static map that shows those points.

---


```{r, echo=T}
library(tidyverse)
library(rnaturalearth)
library(readxl)
wetherspoon_raw <- read_excel("wetherspoon.xlsx")

wetherspoon_raw %>%
  janitor::clean_names() %>% 
  select(1:5)  ->
wetherspoon

wetherspoon
```

```{r wether, include = FALSE, echo=F}
rnaturalearth::ne_countries(
  scale = "medium", returnclass = "sf") %>% 
  filter(name == 'United Kingdom') -> 
  uk

ggplot(data = wetherspoon) + 
  ggplot2::geom_sf(data = uk) + 
  geom_point(aes(x = longitude, 
                   y = latitude, 
                   color = empire_state_burger), 
               alpha = .3) +
    scale_color_viridis_c() + 
    labs(color = "Empire State Burger\nCost (pounds)") +
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5)) + 
    theme(axis.title = element_blank()) 
```


---

`r chunk_reveal("wether")`


---
#Example #2
--

The second example will also draw its shapefiles from the rnaturalearth package. We will also use a dataset containing the defense expenditure of NATO member states. 

We will make a colored map that allows for easy visual comparison of spending levels. This type is called a choropleth map.

---


```{r, echo=T}
library(readxl)
nato_expenditure <- read_excel("nato_expenditure.xlsx")

nato_expenditure
```
---

```{r nato, include = FALSE, echo=F}
rnaturalearth::ne_countries(
  scale = "medium", returnclass = "sf") %>% 
  filter(name %in% c("Albania", "Belgium", "Bulgaria", "Canada", "Croatia", "Czech Republic", "Denmark", "Estonia", "France", "Germany", "Greece", "Hungary", "Iceland", "Italy", "Latvia", "Lithuania", "Luxembourg", "Montenegro", "Netherlands", "Norway", "Poland", "Portugal", "Romania", "Slovakia", "Slovenia", "Spain", "Turkey", "United Kingdom", "United States")) %>% 
  left_join(nato_expenditure,by = "name") ->
  nato_countries

ggplot(data = nato_countries) + 
  ggplot2::geom_sf(data = nato_countries) + 
    aes(fill = spendingPercap) +
    scale_fill_gradient(low = "lightskyblue1", 
                        high = "dodgerblue3") +
  ggplot2::coord_sf(xlim = c(-175, 47.5), 
           ylim = c(23, 85), 
           expand = FALSE) + 
    labs(title = "Defense Expenditure by Share of GDP (%)", 
         subtitle = "NATO Member States", 
         fill = "Share of GDP (%)") + 
    theme_bw() +
    theme(plot.title = element_text(hjust = 0.5), 
          plot.subtitle = element_text(hjust = 0.5)) 
```

`r chunk_reveal("nato", break_type = "auto")`

