---
title: "A collection of COVID-19 Data Visualization"
subtitle: ""
author: "Edited by Gina Reynolds, March 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      beforeInit: "https://platform.twitter.com/widgets.js"
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, 
                      warning = FALSE, comment = "", 
                      cache = F, fig.retina = 3)
library(flipbookr)
library(tidyverse)
```

---

Likely to be discussing this visualization in my Data Viz class, Spring


---


<blockquote class="twitter-tweet"><p lang="en" dir="ltr">NEW on coronavirus: many western countries may soon face Italy’s situation<br><br>Case numbers since outbreaks began in several countries have tracked a ~33% daily rise. This is as true for UK, France, Germany as Italy; the latter is simply further down the path <a href="https://t.co/VcSZISFxzF">https://t.co/VcSZISFxzF</a> <a href="https://t.co/xM6wXuMk4n">pic.twitter.com/xM6wXuMk4n</a></p>&mdash; John Burn-Murdoch (@jburnmurdoch) <a href="https://twitter.com/jburnmurdoch/status/1237737352879112194?ref_src=twsrc%5Etfw">March 11, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


---


<blockquote class="twitter-tweet"><p lang="en" dir="ltr">See the little backgrounds that make the country names in this plot easier to read?<a href="https://twitter.com/jburnmurdoch?ref_src=twsrc%5Etfw">@jburnmurdoch</a> was kind enough to share his <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> code so I could see how it was made (<a href="https://t.co/F7K171s4FR">https://t.co/F7K171s4FR</a>). <br><br>Learned about the shadowtext package by <a href="https://twitter.com/guangchuangyu?ref_src=twsrc%5Etfw">@guangchuangyu</a>: <a href="https://t.co/YKz3h4piKv">https://t.co/YKz3h4piKv</a> <a href="https://t.co/1wYTSUvZuS">https://t.co/1wYTSUvZuS</a></p>&mdash; David Keyes (@dgkeyes) <a href="https://twitter.com/dgkeyes/status/1237812892189093888?ref_src=twsrc%5Etfw">March 11, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

---

# getting data

```{r, eval = F}
url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv"

# if file hasn't already been downloaded, download
if (!file.exists("time_series_19-covid-Confirmed-2020-03-19.csv")){
download.file(url, 
              destfile = 
                "time_series_19-covid-Confirmed-2020-03-19.csv") 
  } # 2020-03-19
```



---


`r chunk_reveal("data_prep")`

```{r data_prep, include=F}
library(tidyverse)

read_csv("time_series_19-covid-Confirmed-2020-03-19.csv") %>%
  gather(key = "date", 
         value = "num_cases",
         5:ncol(.))  %>%
  mutate(date = lubridate::mdy(date)) %>%
  janitor::clean_names() %>% 
  filter(country_region != "Cruise Ship") %>% 
  group_by(country_region, date) %>% 
  summarize(num_cases = sum(num_cases)) ->
data_long

data_long %>% 
  group_by(country_region) %>% 
  summarize(max_cases = max(num_cases)) %>% 
  arrange(-max_cases) %>% 
  slice(1:20) ->
top_20
```

`r chunk_reveal("data_plot")`

```{r data_plot, include=F}
options(scipen = 10)
top_20 %>% 
  left_join(data_long) %>% 
  ggplot() +
  aes(x = date) +
  aes(y = num_cases) +
  geom_line() +
  aes(group = country_region) +
  scale_y_log10() +
  scale_y_log10(breaks = c(1,10,100,1000, 10000, 100000)) +
  scale_y_log10(breaks = c(1,10,100,1000, 10000, 100000),
                limits = c(100, 100000)) ->
g

data_long %>% 
  filter(num_cases > 100) %>% 
  group_by(country_region) %>% 
  mutate(first_data_w_100 = min(date)) %>%
  mutate(days_since_100 = date - first_data_w_100) ->
data_long_since_100

g %+%
  (top_20 %>% 
  left_join(data_long_since_100)) +
  aes(x = days_since_100) +
  theme_minimal() +
  facet_wrap(~ country_region) +
  geom_line(data = . %>%
              mutate(country_region2 = country_region) %>% 
              select(-country_region),
            aes(group = country_region2),
            color = "grey") +
  geom_line(color = "magenta")
```



```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 60%}
```
