---
title: "Data Vizualization of COVID-19"
subtitle: ""
author: "Edited, Gina Reynolds, March 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      beforeInit: "https://platform.twitter.com/widgets.js"
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, 
                      warning = FALSE, comment = "", 
                      cache = T, fig.retina = 3)
library(flipbookr)
library(tidyverse)
```

---

This collection was inspired by John Burns-Murdoch.  He's making some of the most compelling visualizations of this crisis --- at this point, I'd consider his series of COVID-19 figures ast *the* reference point for data vis during this crisis.


---

Here is one of his visualizations.

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">NEW on coronavirus: many western countries may soon face Italy’s situation<br><br>Case numbers since outbreaks began in several countries have tracked a ~33% daily rise. This is as true for UK, France, Germany as Italy; the latter is simply further down the path <a href="https://t.co/VcSZISFxzF">https://t.co/VcSZISFxzF</a> <a href="https://t.co/xM6wXuMk4n">pic.twitter.com/xM6wXuMk4n</a></p>&mdash; John Burn-Murdoch (@jburnmurdoch) <a href="https://twitter.com/jburnmurdoch/status/1237737352879112194?ref_src=twsrc%5Etfw">March 11, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


---

From David Keyes, I learned that he'd shared his code.  

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">See the little backgrounds that make the country names in this plot easier to read?<a href="https://twitter.com/jburnmurdoch?ref_src=twsrc%5Etfw">@jburnmurdoch</a> was kind enough to share his <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> code so I could see how it was made (<a href="https://t.co/F7K171s4FR">https://t.co/F7K171s4FR</a>). <br><br>Learned about the shadowtext package by <a href="https://twitter.com/guangchuangyu?ref_src=twsrc%5Etfw">@guangchuangyu</a>: <a href="https://t.co/YKz3h4piKv">https://t.co/YKz3h4piKv</a> <a href="https://t.co/1wYTSUvZuS">https://t.co/1wYTSUvZuS</a></p>&mdash; David Keyes (@dgkeyes) <a href="https://twitter.com/dgkeyes/status/1237812892189093888?ref_src=twsrc%5Etfw">March 11, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

---


Right after David tweeted this out, I lightly edited John's code to make it apropriate for the flipbook format --- very little adjustment was required.  I'm teaching data viz this spring, so this figure --- and how it's built --- is of interest.  I shared that work with David and John, but John didn't respond, and I've since taken the work down.  

Meanwhile there were a growning number of data visualizations produced by my peers, which much more local focuses --- how were their *homes* being affected.  I've approached them about sharing their work in the "flipbook" format, which communicates the decisions they've made as visualization designers.  

---

In the last couple days, I've created a few plots inspired by John's FT work (i.e. emulating his work --- using my own ggplot know-how instead of his code base) --- and actually going out of the way to try to capture some of the likely decision-making process.  But the work fall well short of the polished product John delivers, with its thematic choices and annotation layers --- and I think these are key to really "selling" the visualization.  Emulation is something that data visualization practicioners do often, but it is important tho think about to what extent is appropriate.  A good discussion is [here](http://www.storytellingwithdata.com/blog/2019/4/1/swdchallenge-emulate), in the context of StoryTellingWithData's "Emulate" Challenge, back in April 2019.  

In what follows, I start with the contributed code (very lightly edited to allow for code walk through.  The emulation of John's work follows.  I think data viz should be understood as a series of insights and decisions.  Walking through many of the insights and decisions help explicitly deliver that message; communicating data viz as process is a message that other visualizers espouse (see the concluding comments of [Chalabi](https://www.lib.ncsu.edu/news/data-visualization-in-the-age-of-%E2%80%9Cfake-news%E2%80%9D)). That's what's being done here, but also with *code* along side the output.  

---

# Alonso Silva

[Alonso Silva](https://scholar.google.com/citations?user=sjlsVQ0AAAAJ&hl=en)  @alonsosilva is a Chilean mathematician residing in France.  He is a senior researcher at at Safran Tech.  He uses the Python programming language and the Matplotlib plotting library to visualize data. He's a family friend. 


He has [blogged](https://medium.com/@alonso.silva/crecimiento-del-n%C3%BAmero-de-casos-confirmados-de-coronavirus-en-chile-bcb1ca4ad9ac) about concern for spread in Chile (in Spanish) on Medium.

---

```{r, child = "alonso_italy_france.Rmd"}

```

---



---


# Peter Higgins

[Dr. Peter Higgins](https://www.med.umich.edu/higginslab/pdrhBio.html) is the director of the IBD program at the University of Michigan, in Ann Arbor. He is also a fan of the R programming language and the ggplot plotting library, which are the tools that he's using for his visualizations about Michigan. I learned about the {covid19us} package from Peter's visualization. His handle is @ibddoctor on Twitter.

---

```{r, child="peter_animation.Rmd"}

```


---




# Sebastián Garrido

Sebastian Garrido is a political scientist at CIDE,with interests in Mexican politics, higher education, data science, and tranparency.  He is an expert of and teaches data visualization, and he uses R and ggplot2 in his work.

His handle is @segasi on Twitter.


---

Placeholder

<blockquote class="twitter-tweet"><p lang="und" dir="ltr"><a href="https://t.co/s982fmmIMw">pic.twitter.com/s982fmmIMw</a></p>&mdash; Sebastián Garrido (@segasi) <a href="https://twitter.com/segasi/status/1243380826395402240?ref_src=twsrc%5Etfw">March 27, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


---

# Emulating the Burns-Murdoch work

This section is meant to give some insight into the decisions process that led to what will probably be the most recognized data visualization from this global tragedy.  Burns-Murdoch uses R and ggplot2 in his work.  

Burns-Murdoch in his [own words on his visualizations] (https://www.ft.com/video/9a72a9d4-8db1-4615-8333-4b73ae3ddff8).

---

```{r, child = "covid_19_2.Rmd"}

```



---


A lot of questions are being raised about Data Viz in the COVID-19 context.

A great summary is in [the Policy Viz Podcast:  Data Viz in the Time of COVID](https://policyviz.com/podcast/episode-170-dataviz-in-the-time-of-covid/).  

[Concerns about attribution]() have also arisen, as this Murdoch's chart has become wildly popular.  



---









```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 60%}
```
