---
title: "Tracking COVID-19"
subtitle: "John Burn-Murdoch's plot for FT"
author: "Edited by Gina Reynolds, March 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      beforeInit: "https://platform.twitter.com/widgets.js"
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



```{r setup, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, 
                      warning = FALSE, comment = "", 
                      cache = TRUE, fig.retina = 3)
library(flipbookr)
```

---

Likely to be discussing this visualization in my Data Viz class, Spring 2020. 


---


<blockquote class="twitter-tweet"><p lang="en" dir="ltr">NEW on coronavirus: many western countries may soon face Italy’s situation<br><br>Case numbers since outbreaks began in several countries have tracked a ~33% daily rise. This is as true for UK, France, Germany as Italy; the latter is simply further down the path <a href="https://t.co/VcSZISFxzF">https://t.co/VcSZISFxzF</a> <a href="https://t.co/xM6wXuMk4n">pic.twitter.com/xM6wXuMk4n</a></p>&mdash; John Burn-Murdoch (@jburnmurdoch) <a href="https://twitter.com/jburnmurdoch/status/1237737352879112194?ref_src=twsrc%5Etfw">March 11, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


---


<blockquote class="twitter-tweet"><p lang="en" dir="ltr">See the little backgrounds that make the country names in this plot easier to read?<a href="https://twitter.com/jburnmurdoch?ref_src=twsrc%5Etfw">@jburnmurdoch</a> was kind enough to share his <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a> code so I could see how it was made (<a href="https://t.co/F7K171s4FR">https://t.co/F7K171s4FR</a>). <br><br>Learned about the shadowtext package by <a href="https://twitter.com/guangchuangyu?ref_src=twsrc%5Etfw">@guangchuangyu</a>: <a href="https://t.co/YKz3h4piKv">https://t.co/YKz3h4piKv</a> <a href="https://t.co/1wYTSUvZuS">https://t.co/1wYTSUvZuS</a></p>&mdash; David Keyes (@dgkeyes) <a href="https://twitter.com/dgkeyes/status/1237812892189093888?ref_src=twsrc%5Etfw">March 11, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>



---

`r chunk_reveal("data_prep")`

```{r data_prep, include=F}
library(tidyverse)
url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_19-covid-Confirmed.csv"
read_csv(url) %>%
  gather(date, cases, 5:ncol(.)) %>%
  mutate(date = as.Date(date, "%m/%d/%y")) %>%
  group_by(country = `Country/Region`, date) %>%
  summarise(cases = sum(cases)) %>%
  filter(country != "Others" & 
           country != "Mainland China") %>%
  bind_rows(
    tibble(country = "Republic of Korea", 
           date = as.Date("2020-03-11"), 
           cases = 7755)
  ) %>%
  group_by(country) %>%
  mutate(days_since_100 = 
           as.numeric(date - min(date[cases >= 100]))) %>%
  ungroup() %>%
  filter(is.finite(days_since_100)) %>% 
  group_by(country) %>%
  mutate(new_cases = cases - cases[days_since_100 == 0]) %>%
  filter(sum(cases >= 100) >= 5) %>%
  filter(cases >= 100) %>% 
  bind_rows(
    tibble(country = "33% daily rise", 
           days_since_100 = 0:18) %>%
      mutate(cases = 100*1.33^days_since_100)
  ) %>%
  ungroup() %>%
  mutate(
    country = country %>% 
      str_replace_all("( SAR)|( \\(.+)|(Republic of )", "")
  ) ->
data
```


---

```{r}
color_assignments <- c("UK" = "#ce3140", "US" = "#EB5E8D", 
                       "Italy" = "black", "France" = "#c2b7af", 
                       "Germany" = "#c2b7af", "Hong Kong" = "#1E8FCC", 
                       "Iran" = "#9dbf57", "Japan" = "#208fce", 
                       "Singapore" = "#1E8FCC", 
                       "Korea" = "#208fce", "Belgium" = "#c2b7af", 
                       "Netherlands" = "#c2b7af", 
                       "Norway" = "#c2b7af", "Spain" = "#c2b7af", 
                       "Sweden" = "#c2b7af", "Switzerland" = "#c2b7af", 
                       "33% daily rise" = "#D9CCC3")
```

---

`r chunk_reveal("plot")`

```{r plot, include = F}
data %>% 
  ggplot() +
  aes(x = days_since_100) +
  aes(y = cases) +
  aes(col = country) +
  geom_hline(yintercept = 100) +
  geom_vline(xintercept = 0) +
  geom_line(size = 0.8) +
  geom_point(pch = 21, size = 1) +
  scale_y_log10(expand = expand_scale(add = c(0,0.1)), 
                breaks = c(100, 200, 500, 1000, 
                           2000, 5000, 10000)) +
  scale_x_continuous(expand = expand_scale(add = c(0,1))) +
  theme_minimal() +
  theme(
    panel.grid.minor = element_blank(),
    legend.position = "none",
    plot.margin = margin(3,15,3,3,"mm")
  ) +
  coord_cartesian(clip = "off") +
  scale_colour_manual(values = color_assignments) +
  shadowtext::geom_shadowtext(
    aes(label = paste0(" ",country)), 
    hjust = 0, vjust = 0, 
    data = . %>% 
      group_by(country) %>% 
      top_n(1, days_since_100), 
    bg.color = "white") +
  labs(x = "Number of days since 100th case", 
       y = "", 
       subtitle = "Total number of cases")
```




```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 60%}
```
