---
title: "The art of flipbooking"
subtitle: "With flipbookr and xaringan"
author: "Gina Reynolds, December 2019"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



<blockquote class="twitter-tweet"><p lang="en" dir="ltr">We&#39;re have no way of measuring the number of cases. This is a chart showing the number of **detected** cases. Is a rapid increase in detected cases a good thing or bad thing? That&#39;s not at all clear. <a href="https://t.co/XYoRV97DDK">https://t.co/XYoRV97DDK</a></p>&mdash; Nate Silver (@NateSilver538) <a href="https://twitter.com/NateSilver538/status/1241117375228977153?ref_src=twsrc%5Etfw">March 20, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


---


<blockquote class="twitter-tweet"><p lang="en" dir="ltr">Today, they reported 189 positives, along with 175 yesterday, as compared with an average of 106 positives per day in the 7 days before that. So, not great on the surface... new cases increased by 70%! But you also have to look at the number of tests.<a href="https://t.co/Goa27xcI1b">https://t.co/Goa27xcI1b</a></p>&mdash; Nate Silver (@NateSilver538) <a href="https://twitter.com/NateSilver538/status/1241113755846684673?ref_src=twsrc%5Etfw">March 20, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

---


https://docs.google.com/spreadsheets/u/2/d/e/2PACX-1vRwAqp96T9sYYq2-i7Tj0pvTf6XVHjDSMIKBdZHXiCGGdNC0ypEU9NbngS8mxea55JuCFuua1MUeOj5/pubhtml#


```{r setup2, include = F}
# This is the recommended set up for flipbooks
# you might think about setting cache to TRUE as you gain practice --- building flipbooks from scracth can be time consuming
knitr::opts_chunk$set(fig.width = 6, message = FALSE, warning = FALSE, comment = "", cache = T, fig.retina = 3)
library(flipbookr)
library(tidyverse)
```

---

`r chunk_reveal("positive_rate")`

```{r positive_rate, include = F}
readxl::read_xlsx("covid_19_tracking_2020_03_20.xlsx") %>%
  janitor::clean_names() %>% 
  data.frame() %>% 
  mutate(date = lubridate::ymd(date)) %>% 
  full_join(tibble(state = state.abb, state.region)) ->
covid_19_data
covid_19_data %>% 
  arrange(date) %>% 
  filter(state == "CO") %>% 
  ggplot() +
  aes(x = total) +
  aes(y = positive) +
  geom_point() +
  geom_line(data = tibble(x = (1:100)^3,
                    y_01 = x * .01,
                    y_03 = x * .03,
                    y_10 = x * .1,
                    y_30 = x * .3,
                    y_100 = x * 1) %>% 
              gather("rate", "y", 2:6),
            aes(x = x, y = y, 
                group = rate),
            color = "blue") +
  aes(color = as.numeric(date)) +
  scale_color_viridis_c() +
  scale_x_log10() +
  scale_y_log10() +
  guides(color = F) ->
g
g %+%
  (covid_19_data %>% filter(state.region == "West")) +
  facet_wrap(~state) 
```



---


```{r, fig.width=10, fig.height=10}
covid_19_data %>% 
  arrange(date) %>% 
  filter(state.region == "West") %>% 
  ggplot() +
  aes(x = date) +
  aes(y = total) +
  geom_area() +
  geom_area(aes(y = positive),
            fill = "magenta") +
  facet_wrap(~state) +
  theme_minimal(base_size = 20)


covid_19_data %>% 
  arrange(date) %>% 
  filter(state.region == "West") %>% 
  ggplot() +
  aes(x = date) +
  aes(y = positive) +
  geom_area() +
  geom_area(aes(y = death),
            fill = "magenta") +
  facet_wrap(~state) +
  theme_minimal(base_size = 20)


covid_19_data %>% 
  arrange(date) %>% 
  filter(state.region == "West") %>% 
  ggplot() +
  aes(x = date) +
  aes(y = death) +
  geom_area() +
  facet_wrap(~state) +
  theme_minimal(base_size = 20) 


```




---



```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 60%}
```
