---
title: "Klaus Schulte's ggplots, uses Claus Wilke's new ggtext"
subtitle: "made with {flipbookr}"
author: "Edited by Gina Reynolds, Feb 2020"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: [default, hygge, ninjutsu]
    nature:
      beforeInit: "https://platform.twitter.com/widgets.js"
      ratio: 16:10
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---



My old twitter friend from #MakeoverMonday tried ggplot for the first time a few weeks ago.  He figured out tons of things about ggplot2, and even used Claus Wilke's new ggtext package, and the new ggplot theme option `plot.title.position`.  Way to go Klaus!  And way to go Claus!  




<blockquote class="twitter-tweet"><p lang="en" dir="ltr">Getting started with <a href="https://twitter.com/hashtag/Rstats?src=hash&amp;ref_src=twsrc%5Etfw">#Rstats</a>, <a href="https://twitter.com/hashtag/rstudio?src=hash&amp;ref_src=twsrc%5Etfw">#rstudio</a> and <a href="https://twitter.com/hashtag/ggplot2?src=hash&amp;ref_src=twsrc%5Etfw">#ggplot2</a> was my project of the week. My first chart was a lot more work than expected and is inspired by <a href="https://twitter.com/barnabasmarkus?ref_src=twsrc%5Etfw">@barnabasmarkus</a> <a href="https://twitter.com/hashtag/makeovermonday?src=hash&amp;ref_src=twsrc%5Etfw">#makeovermonday</a> from week 1. Thanks <a href="https://twitter.com/EvaMaeRey?ref_src=twsrc%5Etfw">@EvaMaeRey</a> for support, my code is probably lousy, but I will build up on this! <a href="https://t.co/2dggPFrwrX">pic.twitter.com/2dggPFrwrX</a></p>&mdash; Klaus Schulte (@ProfDrKSchulte) <a href="https://twitter.com/ProfDrKSchulte/status/1222160205351464960?ref_src=twsrc%5Etfw">January 28, 2020</a></blockquote> <script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>


---

```{r}
library("httr")
library("readxl")
df <- readxl::read_excel("sports_data.xlsx")
knitr::opts_chunk$set(cache = T, comment = "")
```

---

```{r}
library(tidyverse)
```

---

`r flipbookr::chunk_reveal("wrangle")`

```{r wrangle, include = F}
df %>% 
  janitor::clean_names() %>%
  arrange(-x2017) %>% 
  select(1:2) %>% 
  mutate(sport_dummy = sport != "None") %>% 
  slice(1:10) %>% 
ggplot() +
  aes(x = sport) +
  aes(y = x2017) + 
  geom_col(width = 0.6) + 
  aes(fill = sport_dummy) +
  scale_fill_manual(values =
                      c("#F11B59", "#8d8d8d")) + 
  coord_flip() + 
  aes(x = reorder(sport, x2017)) +
  geom_text(aes(label = x2017),
            nudge_y = 1) +
  aes(color = sport_dummy) + 
  scale_color_manual(values = 
                       c("#F11B59", "#8d8d8d")) + 
  theme_minimal() +
  theme(legend.position = "none", #no legend
        plot.margin = margin(35, 35, 10, 35)) + 
  labs(x = NULL) + 
  labs(y = NULL) ->
g
```


---

`r flipbookr::chunk_reveal("ggtext")`

```{r ggtext, include = F}
# now all the ggtext goodness
g +
  labs(caption = paste0(
    "#makeovermonday 01/2020 | ",
    "@ProfDrKSchulte | Inspired by: @barnabasmarkus")
    ) +
  theme(plot.caption = ggtext::element_textbox_simple(
      size = 10,
      lineheight = 1,
      padding = margin(10, 10, 10, 10),
      margin = margin(10, 0, 10, 0),
      fill = "#F5F5F5",
      halign = 0.5,
      valign = 0.5)) +
    labs(title = paste0(
      "<b><span style = 'font-size:20pt'>15% of",
      "<span style = 'color:#F11B59;'> ",
      "**Americans Have no Favorite Spectator Sport**",
      "</span></span></b>",
      "<br><b><span style = 'font-size:14pt'>",
      "Share of Answers to **Gallup** question ",
      "*'What is your favorite sport to watch?'* ",
      "in 20017",
      "</span></b><br>")
       ) +
    theme(plot.title = ggtext::element_textbox_simple(
      size = 13,
      face = NULL,
      lineheight = 1.75,
      padding = margin(5, 5, 0, 5),
      margin = margin(0, 0, 0, 0),
      fill = "white")) +
  theme(plot.title.position = "plot")

```

---

Demoing some of what Claus Demoed

---

`r flipbookr::chunk_reveal("claus_ggtext")`

```{r claus_ggtext, include = F}
library(ggtext)
iris %>% 
  mutate(my_var = 
           glue::glue("*{Species}* (Species)")) %>% 
  mutate(my_color = 
           case_when(Species == "setosa" ~ "#009456",
                     Species == "versicolor" ~ "#827139",
                     Species == "virginica" ~ "#938290")) %>% 
  select(my_var, my_color, everything()) ->
prepped

prepped %>% 
  ggplot() + 
  aes(x = my_var) +
  aes(y = Sepal.Width) +
  geom_boxplot() +
  geom_jitter(width = .2,
              height = 0,
              alpha = .5) +
  aes(fill = my_color) +
  scale_fill_identity() +
  coord_flip() + 
  ggtitle("hello *goodbye*") +
  theme(plot.title = 
          element_markdown()) +
  theme(axis.text.y = 
          element_markdown()) +
  geom_richtext(data = . %>% 
                  group_by(Species, my_var, my_color) %>% 
                  summarise(the_mean = mean(Sepal.Width)),
                aes(y = the_mean, 
                    label = glue::glue(
                      "*mean* = {the_mean} <br>*mean*<sup>2</sup> = {the_mean^2}"))
                ) ->
g

g %+% 
  (prepped %>% 
     mutate(my_var = 
              glue::glue("<i style='color:{my_color}'>{Species}</i> ({Species})")))
```




```{css, eval = TRUE, echo = FALSE}
.remark-code{line-height: 1.5; font-size: 55%}
```

